Ethnicity sensitive author disambiguation using semi-supervised learning

### 2015-08-25 paper/paper.tex
title
save

\title{Ethnicity sensitive author dissambiguation using semi-supervised learning}

@@

Efforts and solutions to author disambiguation have been proposed from various
communities \citep{liu2014author}. On the one hand, libraries have maintained
authorship control through manual curation, either in a centralized way by
hiring collaborators or through developing services that invite authors to register
their publications themselves (\eg, Google Scholar, and Inspire-HEP).
Recent efforts to create persistent digital identifiers assigned to researchers (\eg, ORCID, and ResearchedID), 
with the objective to embed these identifiers in the submission workflow of publishers
or repositories (\eg, Elsevier, arXiv, Inspire-HEP), would univocally solve any disambiguation issue.
With the large cost of centralized manual authorship control, or until crowdsourced solutions are more 
widely adopted, the impact of these efforts are unfortunately limited by the efficiency, motivation
and integrity of their active contributors. Similarly, the success of persistent digital identifier efforts 
is conditioned to a large and ubiquitous adoption by both researchers and publishers.
For these reasons, fully automated machine learning-based methods have been proposed during the past decade to
provide immediate, less costly, and satisfactory solutions to author
disambiguation. In this work, our goal is to explore and demonstrate how both
approaches can coexist and benefit from each other.
In particular, we study how labeled data obtained through manual curation (either centralized or
crowdsourced) can be exploited (i) to learn an accurate classifier for
identifying coreferring authors, and (ii) to guide the clustering of scientific
publications by distinct authors in a semi-supervised way.
Our analysis of parameters and features of this large dataset reveal that the general pipeline
commonly used in existing solutions is an effective approach for author disambiguation.
Additionally, we propose alternative strategies for blocking based on the
phonetization of author names to increase recall.
We also propose ethnicity-sensitive features for learning a linkage function,
thereby tailoring author disambiguation to non-Western author names whenever necessary.

@@

Training data is usually not easily available, therefore unsupervised approaches propose
use of domain-specific, manually designed linkage functions tailored towards author
disambiguation \citep{malin2005unsupervised,mcrae2006also,song2007efficient,
soler2007separating, kang2009co,fan2011graph,schulz2014exploiting}.
These approaches have the advantage of not requiring hand-labeled data, but generally do
not perform as well as supervised approaches.
To reconcile both worlds, semi-supervised methods make use of small, manually verified, clusters of
publications and/or high-precision domain-specific rules to build a training
set of pairs of publications, from which a linkage function is then built using supervised learning
\citep{ferreira2010effective,torvik2009author,levin2012citation}.

@@

In this context, we position the work in this paper as a
semi-supervised solution for author disambiguation, with the significant
advantage of having a very large collection of more than 1 million crowdsourced annotations
of publications whose true authors are identified.
The extent and coverage of this data allows us to revisit, validate and nuance previous 
findings regarding supervised learning of linkage functions, and to better explore strategies
for semi-supervised clustering.
Furthermore, by releasing our data in to the public domain, we hope to provide a benchmark on 
which further research on author disambiguation and related topics can be built.

@@

The simplest and most common strategy for blocking, referred to hereon in as \emph{SFI},
groups signatures together if they share the same surname(s) and the same first
given name initial (\eg, \emph{SFI}("Doe, John") $==$ "Doe, J").
Despite satisfactory performance, there are several cases where this simple strategy fails to cluster related pairs of signatures together, including:

@@

\textit{Supervised classification.} The second part of the algorithm is the
automatic construction of a pair-wise linkage function between signatures for use 
during the clustering step which grouping all signatures from the same author.

@@

Second, properly constructing a training set of positive and negative pairs of
signatures from which to learn a linkage function yields a significant
improvement.
A random sampling of positive and negative pairs, without taking
blocking into account, significantly impacts the overall performance
($0.9711$). When pairs are drawn only from blocks, performance increases to
($0.9786$), which confirms our intuition that $d$ should be built only from
pairs it will be used to eventually cluster. Finally, making the classification
problem more difficult by oversampling complex cases proves to be relevant,
by further improving the disambiguation results ($0.9830$).

>>>>

\title{Ethnicity sensitive author disambiguation using semi-supervised learning}

@@

Efforts and solutions to author disambiguation have been proposed from various
communities \citep{liu2014author}. On the one hand, libraries have maintained
authorship control through manual curation, either in a centralized way by
hiring professional collaborators or through developing services that invite authors to register
their publications themselves (\eg, Google Scholar or Inspire-HEP).
Recent efforts to create persistent digital identifiers assigned to researchers (\eg, ORCID or ResearchedID),
with the objective to embed these identifiers in the submission workflow of publishers
or repositories (\eg, Elsevier, arXiv or Inspire-HEP), would univocally solve any disambiguation issue.
With the large cost of centralized manual authorship control, or until crowdsourced solutions are more
widely adopted, the impact of these efforts are unfortunately limited by the efficiency, motivation
and integrity of their active contributors. Similarly, the success of persistent digital identifier efforts
is conditioned to a large and ubiquitous adoption by both researchers and publishers.
For these reasons, fully automated machine learning-based methods have been proposed during the past decade to
provide immediate, less costly, and satisfactory solutions to author
disambiguation. In this work, our goal is to explore and demonstrate how both
approaches can coexist and benefit from each other.
In particular, we study how labeled data obtained through manual curation (either centralized or
crowdsourced) can be exploited (i) to learn an accurate classifier for
identifying coreferring authors, and (ii) to guide the clustering of scientific
publications by distinct authors in a semi-supervised way.
Our analysis of parameters and features of this large dataset reveal that the general pipeline
commonly used in existing solutions is an effective approach for author disambiguation.
Additionally, we propose alternative strategies for blocking based on the
phonetization of author names to increase recall.
We also propose ethnicity-sensitive features for learning a linkage function,
thereby tailoring author disambiguation to non-Western author names whenever necessary.

@@

Training data is usually not easily available, therefore unsupervised approaches propose
the use of domain-specific, manually designed, linkage functions tailored towards author
disambiguation \citep{malin2005unsupervised,mcrae2006also,song2007efficient,
soler2007separating, kang2009co,fan2011graph,schulz2014exploiting}.
These approaches have the advantage of not requiring hand-labeled data, but generally do
not perform as well as supervised approaches.
To reconcile both worlds, semi-supervised methods make use of small, manually verified, clusters of
publications and/or high-precision domain-specific rules to build a training
set of pairs of publications, from which a linkage function is then built using supervised learning
\citep{ferreira2010effective,torvik2009author,levin2012citation}.

@@

In this context, we position the work in this paper as a
semi-supervised solution for author disambiguation, with the significant
advantage of having a very large collection of more than 1 million crowdsourced annotations
of publications whose true authors are identified.
The extent and coverage of this data allows us to revisit, validate and nuance previous
findings regarding supervised learning of linkage functions, and to better explore strategies
for semi-supervised clustering.
Furthermore, by releasing our data in the public domain, we hope to provide a benchmark on
which further research on author disambiguation and related topics can be built.

@@

The simplest and most common strategy for blocking, referred to hereon as \emph{SFI},
groups signatures together if they share the same surname(s) and the same first
given name initial (\eg, \emph{SFI}("Doe, John") $==$ "Doe, J").
Despite satisfactory performance, there are several cases where this simple strategy fails to cluster related pairs of signatures together, including:

@@

\textit{Supervised classification.} The second part of the algorithm is the
automatic construction of a pair-wise linkage function between signatures for use
during the clustering step which groups all signatures from the same author.

@@

Second, properly constructing a training set of positive and negative pairs of
signatures from which to learn a linkage function yields a significant
improvement.
A random sampling of positive and negative pairs, without taking
blocking into account, significantly impacts the overall performance
($0.9711$). When pairs are drawn only from blocks, performance increases
($0.9786$), which confirms our intuition that $d$ should be built only from
pairs it will be used to eventually cluster. Finally, making the classification
problem more difficult by oversampling complex cases proves to be relevant,
by further improving the disambiguation results ($0.9830$).
